Manuscript Heat Map Script

### 2025-05-06 ABP_heatmap_for_Cece_20241018.Rmd
Revising scripts to include all figures

---
title: "Cece Sed Traps Manuscript Heat Map"
author: "Adrienne Breef-Pilz"
date: "2024-10-18"
output: html_document
---

@@

 EDI_CTD <-
  read_csv("https://pasta.lternet.edu/package/data/eml/edi/200/14/0432a298a90b2b662f26c46071f66b8a")%>%
  #read_csv("./../../Raw_CTD/ctd_L1.csv")|>
   filter(DateTime>ymd_hms("2018-01-01 00:00:00") & DateTime<ymd_hms("2023-12-31 00:00:00"))%>%
   filter(Reservoir=="BVR"|Reservoir=="FCR")%>%
   filter(Site==50)%>%
   select(-starts_with("Flag"))%>%
  mutate(Year=year(DateTime))%>%
  filter(Depth_m>0)

@@

### Still working on this

@@

```{r Need to add in handhelp pH obs}

ph <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/198/12/e0181372c6d4cbc66eca4644f8385470" )|>
  #read_csv("./../../../DataAlreadyUploadedToEDI/EDIProductionFiles/MakeEMLYSI_PAR_secchi/2023/YSI_PAR_profiles_2013_2023.csv")|>
  filter(DateTime>as.Date("2018-01-01"))|>
  filter(Reservoir=="BVR"|Reservoir=="FCR")|>
   filter(Site==50)|>
  select(Reservoir, DateTime, Depth_m, pH)|>
  drop_na(pH)|>
  mutate(Year=year(DateTime))

# bind with the CTD

EDI_CTD_ph <- dplyr::bind_rows(EDI_CTD,ph)

CTD_casts_ph <- ph|>
  select(Reservoir, DateTime, Year)|>
  mutate(DateTime = as.Date(DateTime))|>
  unique()%>%
  dplyr::bind_rows(. , CTD_casts)

# GEt the month and date for easier graphing. This is not the correct date.   
  CTD_casts_ph$MDay <- format(CTD_casts_ph$DateTime,"%m-%d")
  CTD_casts_ph$MDay <- as.Date(CTD_casts_ph$MDay, "%m-%d")
```

```{r Make pH plot}

##Need to fix this

 EDI_CTD_pH <- EDI_CTD|>
   select(Reservoir, DateTime, Year, Depth_m, pH)%>%
  bind_rows(.,ph)

wer <- data.frame(
  data= c(rep("EDI_CTD_pH",12)),
  var = c(rep("pH",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

make_heat_map(
  data= temp_df,
  sample_dates = CTD_casts_ph,
  turnovers = turnover,
  y_lim = c(5,10),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "pH",
  output_file = "figures/SedTraps_CTD_pH_2018_2023.jpeg"
)

```{r metals plotting}

@@

# change the plots dimensions  
#jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "gray80") + #to mark cast dates 
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="gray80", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0),
                  breaks = c(0.0, 2.5, 5.0, 7.5)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank()) +
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette,
                       values=rescale(c(0,5,20)))+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)
  # labs(color = "")+
  # geom_point(data = sensor_depth_plot, 
  #            aes(x = Date, y = Sensor_depth), 
  #            shape = 23, fill = "black", color = "white", size  =2.5)+
  # geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  # scale_linetype_manual(values = c("solid","84"),
  #                       limits = c("Thermocline","Outtake pipe"),  
  #                       name= NULL)+

@@

```{r}
metals <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/455/8/9c8c61b003923f4f03ebfe55cea8bbfd")

@@

>>>>

---
title: "Manuscript Heat Map Script"
author: "Adrienne Breef-Pilz"
date: "2024-10-18"
output: html_document
---

Includes script for making Figures 2-5 and S2. Script written by ABP and updated by CEW. Edits include code chunks written by CEB.

@@

 EDI_CTD <-
  read_csv("https://pasta.lternet.edu/package/data/eml/edi/200/14/0432a298a90b2b662f26c46071f66b8a")%>%
   filter(DateTime>ymd_hms("2018-01-01 00:00:00") & DateTime<ymd_hms("2023-12-31 00:00:00"))%>%
   filter(Reservoir=="BVR"|Reservoir=="FCR")%>%
   filter(Site==50)%>%
   select(-starts_with("Flag"))%>%
  mutate(Year=year(DateTime))%>%
  filter(Depth_m>0)

@@

@@

```{r metals plotting script}

@@

# change the plots dimensions  
#jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "gray80") + #to mark cast dates 
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="gray80", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0),
                  breaks = c(0.0, 2.5, 5.0, 7.5)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank()) +
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette,
                       values=rescale(c(0,5,20)))+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)

@@

```{r read in metals data}
metals <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/455/8/9c8c61b003923f4f03ebfe55cea8bbfd")

@@

## Soluble:Total Metal Heat Maps (Figure S2)

```{r S:T metals plotting}

make_metals_ratio_heat_map <- function(
    data,
    sample_dates,
    turnovers, 
    y_lim,
    custom_heatmap_palette,
    y_units){

# change the plots dimensions  
#jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "gray60") + #to mark cast dates 
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="gray60", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0),
                  breaks = c(0.0, 2.5, 5.0, 7.5)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank()) +
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette)+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)

```{r Make S:T Fe plot}
wer <- data.frame(
  data= c(rep("metratio",12)),
  var = c(rep("Fe_ratio",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

p5 <- make_metals_ratio_heat_map(
  data= temp_df,
  sample_dates = metDates,
  turnovers = turnover, 
  y_lim = c(0,1),
  custom_heatmap_palette = blue2red(60),
  y_units = "Soluble:Total"
) + ggtitle('Iron')

```{r Make S:T Mn plot}
wer <- data.frame(
  data= c(rep("metratio",12)),
  var = c(rep("Mn_ratio",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

p6 <- make_metals_ratio_heat_map(
  data= temp_df,
  sample_dates = metDates,
  turnovers = turnover, 
  y_lim = c(0,1),
  custom_heatmap_palette = blue2red(60),
  y_units = "Soluble:Total"
) + ggtitle('Manganese')

```{r}
STPlot <- ggarrange(p5, p6, ncol = 1, labels = c('A', 'B'), common.legend = TRUE,
                    legend = 'bottom')
STPlot
ggsave(STPlot, filename = "figures/SedTraps_STRatio_2018_2023.png", height = 6, width = 8, units = 'in')
```
